Browsernative Microservices

### 2017-02-09 build/paper.tex
pdf/tex build

The spirit of freedom can't be applied to \emph{browsernative
microservices} as the browser and its underlying DOM will be the
limitation factor. Still there are some decentralization aspects around
creating microservices for the browser worth to mention.

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for
different reasons. Transpiling languages to JS as target language isn't
exotic anymore. Languages like TypeScript, ClojureScript or PureScript
compile to JS even exclusively. Once web components hit a critical mass
there will be most likely some library support or foreign function
interfaces towards ES6 modules (which are mandatory for the new
specifications). With the rise of WebAssembly, a new low-level
programming language for the browser, the determination on JS will
hypothetical deteriorate and new quasi native languages for the web
might gain traction.

Another more real life decentralization aspect derives from the easiness
of deployment of \emph{browsernative microservices} which makes them a
good fit for open source. They are small, leightweight and can be tested
live in the browser without any additional tooling.
\href{https://webcomponents.org}{Webcomponents.org} for example is a
registry for ready-to-use components of every scale where even Google
shares a lot of their material design elements.

Decentralization will be achieved just because of the encapsulation of
the service. Grouping together HTML, JS and CSS Code in a safe,
sandboxed environment exposes the possibility to build more cohesive and
understandable services. In the typical global nature of web development
those three pillars are separated. This circumstance left the developer
switching back and forth between code bases developing a tricky (and
sometimes biased) way to glue related parts together.

@@

@@

Chapter about progressive enhancement

@@

Shift of paradigms / BFF / Platform agnostic / changes in infrastructure
like APIs Databanks / Deployment

The ideas transcending from the microservice approach offers plenty of
choices and decisions how to proceed with designing a program or to
structure a process.

@@

JS runs in a single-threaded language environment. Every script in the
browser environment, from handling UI events to query and process larget
amounts of API data and manipulating the DOM, runs on the same
thread{[}15{]}. Putting a lot of work to the single main thread can slow
down the web service significantly. From time to time scripts can block
or fail to whatever reason which leads to a frozen UI on the users side.
A worker can overcome the bottleneck of the single-threaded nature with
spawning a new \textbf{background thread}. Most of the browsers work can
be leveraged to this new thread. Todays web architectures aims to
leverage an increasing amout of proccessing to the client to avoid
time-consuming roundtrips especially in mobiles networks.\footnote{Latency
  numbers: https://gist.github.com/jboner/2841832} While many native
APIs like \texttt{fetch} work seamlessly in the new thread, a worker has
no access to the DOM at all.

@@

Having no access to the DOM can be seen as hinderance,but i

Macroperspektive / Composition

\subsubsection{Assumptions about custom
elements}\label{assumptions-about-custom-elements}

Apart from the spec'd perspective there is mental model a webdeveloper
might

Creating and using webcomponents might require a new mental model how do
design a

containers

@@

>>>>

The spirit of freedom can't be applied to \emph{browsernative
microservices} as the browser and its underlying DOM will be the
limitation factor. Talking about the browser, a reader might be tempted
to narrowly thinking about the VIEW layer only - which is not true
anymore. In the recent years the browser engines grows to a kind of
virtual machine: there are connectors to build-in databases,
multithreading support, an ever-growing JS build-ins and even push
notifications. So-called \emph{Progressive Web Apps\footnote{\href{https://developers.google.com/web/progressive-web-apps/}{Progressive
  Web Apps}}}, a bunch of criteria for building good browserapps, can
achieve a similar look and feel like native apps.

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for various
reasons. Transpiling languages to JS as target language isn't exotic
anymore. Languages like TypeScript, ClojureScript or PureScript compile
to JS even exclusively. Once web components hit a critical mass there
will be most likely some library support or foreign function interfaces
towards ES6 modules (which are mandatory for the new specifications).
With the rise of WebAssembly, a new low-level programming language for
the browser, the determination on JS will hypothetical deteriorate and
new quasi native languages for the web might gain traction.

Another more real life decentralization aspect derives from the easiness
of deployment in a safe, sandboxed environment. No hidden overheads or
required build tools. Thats why \emph{browsernative microservices} are
well suited for sharing and open-sourcing. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
is a registry for ready-to-use components of every scale where for
example Google shares a lot of their material design elements.

\subsection{Decentralized Data
Management}\label{decentralized-data-management}

Data Management in Microservice follows the same modular philosophy like
the conceptual implementation. As mentioned earlier different bounded
contexts make different assumptions of the underlying model. A
\emph{browsernative microservices} takes this thought even further and
expands it to the variety of devices. Decentralized decisions about
conceptual models demand for decentralized data storage
decisions.{[}4{]} Since network roundtrips are costly it is a good
advice to only query as much data as needed and cache as much as
possible.

``Microservices prefer letting each service manage its own
database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
emphasizes this pragmatism on APIs in a conference talk. At ING the
frontend demands drive the APIs, APIs are produces automatically and not
even Issa knows how many APIs exists.{[}9{]} They are using a pattern
called \textbf{backend for frontends} allowing the team working on a
given UI also handle the server-side components.{[}3, p. 72{]} The
perception on backend might change therefore from embodying logic to a
data-silo only.

To see this pattern in the field a reader might have a look at Facebooks
GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
query language for the frontend. The backend solely replies on the
frontend needs. Another well documented \sout{even more powerful}
approach in the field is Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
where parts of the database will be reflected to the client. A so-called
Transactor ensures ACID compliance.

The upcoming microservice example assumes a generic build-in API next to
the build-in frontend components. Instead of gluing frontend and backend
together on runtime, the microservice is designed holistically
containing both ends. To reduce network calls especially for mobile
devices it is a good advice to cache data via a global Service Worker.
Revamping offline capabilities even further data can be stored in a
browser based database like PouchDB. For the sake of simplicity data
management won't be conducted into depth throughout this paper.

@@

In the global nature of web development the developer stage couldn't't
completely decoupled from the production environment. This circumstance
left developers switching back and forth between files developing tricky
(and more often biased) ways to glue related parts together. With scoped
components, code blocks can be developed more sane and conveyed into
production without headache. Even more, in the scripted environment of
the browser this can be implemented rapidly and continuous.

Previously mentioned Ben Issa, described the ING workflow pretty simple.
Every component deserves a own GIT repo containing

\begin{itemize}
\tightlist
\item
  Internationalization conformity (i18n)
\item
  Accessibility conformity (a11y)
\item
  Tests for the component
\item
  Demos of the component
\item
  Blueprints to mock the one to one APIs
\item
  Docs
\end{itemize}

Every check-in is handled as release candidate and can be independently
deployed by a fully automated deployment machinery.{[}9{]} While this
example is seems opinionated, it gives a sense of a mature component
build for the web. Due to an exhaustive amount of testing and deployment
tools for the browser an automated infrastructure shouldn't be a
problem.

@@

In theory a microservice is designed with a lot of emphasizes on
real-time monitoring for both the architectural elements and business
relevant metrics.{[}4{]} Due to the modular structure weak points can
occur in the orchestration or the services. Testing and automation is a
feasible task but failures occur within browsers anyway. For example,
legacy browsers remain a problem for enhancing websites with new
technologies.

Regarding the evolution of the web, the ``next billion'' internet users
most likely using Android, have decent specs mobile phones, use an
evergreen browser but won't have a reliable internet connection.{[}10{]}
While \emph{Progressive Enhancement} was once coined on the principle to
build websites both for Browsers with JS and without, the new
\emph{Progressive Enhancement} becomes more an ``offline first''
principle. A \emph{browsernative microservice} therefore not only tries
to cache data as much as possible, it should also bring in a lot of
program logic as described in the previous chapters.

Working in a JS heavy infrastructure demands for optimazation to avoid
unexpected side-effects like the \emph{flash of unstyled content
(FOUC)}. Googles Polymer propagates the a general-purpose \textbf{PRLP
pattern}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
  pattern}}:

\begin{itemize}
\tightlist
\item
  Push critical resources for the initial route
\item
  Render initial route
\item
  Pre-cache remaining routes
\item
  Lazy-load and create remaining routes on demand
\end{itemize}

Following this pattern a critical resource can detect browser
functionalities beforehand and switch to a \textbf{polyfill} instead of
the latest browser optimized version. After the initial paint, critical
resources like top-level microservices or other app logic can be loaded
and registered.

@@

Using modular web architecture isn't a perceptually new approach.

Angular, React

Interopt

Freedom of choice

@@

JS runs in a single-threaded language environment. Every script in the
browser environment, from handling UI events to query and process larget
amounts of API data and manipulating the DOM, runs on the same
thread{[}16{]}. Putting a lot of work to the single main thread can slow
down the web service significantly. From time to time scripts can block
or fail for whatever reason which leads to a frozen UI on the users
side. A worker can overcome the bottleneck of the single-threaded nature
with spawning new \textbf{background threads}. Todays web architectures
aims to leverage an increasing amout of processing to the client to
avoid time-consuming roundtrips especially in mobiles
networks.\footnote{Latency numbers:
  https://gist.github.com/jboner/2841832} While many native APIs like
\texttt{fetch} work seamlessly in the new thread, a worker has no access
to the DOM at all.

@@

@@

After grasping the principles and some technical background the paper
should illustrate a practical example to \emph{browsernative
microservices}. Googles library Polymer is a good start to learn about
web components. They even offer a command line tool for creating a
polymer skeleton. One of their most famous proof of concept is the
so-called \href{https://shop.polymer-project.org/}{Polymer Shop} which
is a fully-fledged online shop nested within a single custom element
\texttt{\textless{}shop-app\textgreater{}}. It has some elements for
routing, managing view, service worker caching, theming, etc. The whole
shop runs as a single application only fetching and sending resources
but never reloading altogether. Let's assume we work in a sales
engineering team of the Polymer Shop and need to rebuild checkout
process.

A usual checkout has 3 to 4 steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Sign in or Sign up
\item
  Shipping details
\item
  Payment details
\item
  Review and place order
\end{enumerate}

Translated into a raw custom element HTML structure, the microservice
might light like:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{<shop-checkout>}
  \KeywordTok{<sign-in></sign-in>}
  \KeywordTok{<shipping-details></shipping-details>}
  \KeywordTok{<payment-details></payment-details>}
  \KeywordTok{<place-order></place-order>}
\KeywordTok{</shop-checkout>}
\end{Highlighting}
\end{Shaded}

Pretty straight forward.

http://www.pocketjavascript.com/blog/2015/11/23/introducing-pokedex-org

bare minimum of centralized management of these services

\section{Thinking further}\label{thinking-further}

endgeräte werden besser

multicore

query languages from handheld
